---
title: "Website Analytics - Arctic Data Center"
date: "Last Updated May 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Arctic Data Center Website Metrics

In addition to tracking Twitter metrics, I'm also interested in tracking traffic to arcticdata.io and different pages within that domain. 

# Data Collection - Summary

I collected data manually from Google Analytics once a month near the first of each month and added it to the folder Website Metrics. This summary file contains:

* **month**: month of Tweet information
* **year**: year of Tweet information
* **date**: formatted the month as the first day of the months o that the date would format nicely for as.Date
* **pubs_view**: number of pageviews for the publications page - https://arcticdata.io/publications/
* **home_view**: number of pageviews for the home page - https://arcticdata.io/
* **catalog_view**: number of pageviews for the catalog page - https://arcticdata.io/catalog
* **catalogdata_view**: number of pageviews for the catalog/data page - https://arcticdata.io/catalog/data
    * This is the page that people are directed to when they go back to the search from the Metacat site rather than the Wordpress site.
* **data_view**: number of pageviews for the data page
    * In October 2018, the catalog page was created - before that time, the data page URL was https://arcticdata.io/data. On and before October 2018, these are the pageviews for https://arcticdata.io/data. On and after November 2018, this variable is the sum of the catalog_view and catalogdata_view.
* **submit_view**: number of pageviews for the data submission page - https://arcticdata.io/catalog/submit
* **site_view**: Number of pageviews for the entire arcticdata.io domain.
    * These are all the total pageview count for the associated pages, which is a total count and doesn't exclude repeat visitors or Arctic Data Center staff views.

## Data Analysis

```{r load-packages, echo=FALSE}
library(readr)
library(knitr)
library(ggplot2)

```

```{r raw-data, echo=FALSE,results=FALSE,message=FALSE}
website_summary_metrics <- read_csv("~/Documents/Twitter Analytics/Raw Data/website_summary_metrics.csv")
website_summary_metrics$Date <-as.Date(website_summary_metrics$Date) 

website_summary_metrics$Pub_Percent<-(website_summary_metrics$Pubs_View/website_summary_metrics$Site_View)*100
website_summary_metrics$Data_Percent<-(website_summary_metrics$Data_View/website_summary_metrics$Site_View)*100
website_summary_metrics$Submit_Percent<-(website_summary_metrics$Submit_View/website_summary_metrics$Site_View)*100
```

```{r display-table, echo=FALSE,results=FALSE,message=FALSE}
knitr::kable(website_summary_metrics, caption = "Website Metrics Table")
```

## Graphing Data

```{r creating-custom-theme, echo=FALSE,results=FALSE,message=FALSE}
theme_ADC<- function() {
  theme_bw(base_size=12,base_family="Helvetica") %+replace%
    theme(
      plot.title=element_text(size=20, face="bold",margin=margin(10,0,10,0),color="#1D244F"),
        axis.text.x = element_text(angle=50, size=10, vjust=0.5, color="#1D244F"),
        axis.text.y = element_text(size=10, color="#1D244F"),
        axis.title.x = element_text(color="#1D244F",vjust=-.5,size=14),
        axis.title.y = element_text(color="#1D244F",angle=90,vjust=.5,size=14),
        panel.background=element_rect(fill="white"),
        axis.line = element_line(color="#1D244F"),
      panel.grid.major = element_line(colour = "white", size = 0.2), 
    panel.grid.minor = element_line(colour = "white", size = 0.5),
    )
}
```

```{r monthly-views, echo=FALSE,results=FALSE,message=FALSE}
monthlyviews<-ggplot(data = website_summary_metrics, aes(x=Date, y=Home_View)) +
  geom_point(color="#156760",size=2.5)+
  labs(title='Monthly Page Views Over Time',
       x="Date",
       y="Page Views", hjust=.5)

monthlyviews+theme_ADC()
```

```{r searcher-views, echo=FALSE}

lm_eqn <- function(website_summary_metrics){
    m <- lm(Data_View ~ Date, website_summary_metrics);
    eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2, 
         list(a = format(unname(coef(m)[1]), digits = 2),
              b = format(unname(coef(m)[2]), digits = 2),
             r2 = format(summary(m)$r.squared, digits = 3)))
    as.character(as.expression(eq));
}

searchers<-ggplot(data = website_summary_metrics, aes(x=Date, y=Data_View)) +
  geom_point(color="#156760",size=2.5)+
  labs(title='Monthly Search Catalog Page Views Over Time',
       x="Date",
       y="Search Catalog Page Views")+
  stat_smooth(method="lm", se=FALSE, color="#B4E6EA")

searchers+theme_ADC() + 
  geom_text(data=website_summary_metrics[40, ], y=1100, label = lm_eqn(website_summary_metrics), parse = TRUE)



```


```{r submitters, echo=FALSE}
submitters<-ggplot(data = website_summary_metrics, aes(x=Date, y=Submit_View)) +
  geom_point(color="#156760",size=2.5)+
    labs(title='Monthly Sumbit Page Visits Over Time',
       x="Date",
       y="Submit Page Visits")

submitters+theme_ADC()
```


```{r publications, echo=FALSE}
publications<-ggplot(data = website_summary_metrics, aes(x=Date, y=Pubs_View)) +
  geom_point(color="#156760",size=2.5)+
  labs(title='Monthly Publications Page Visits Over Time',
       x="Date",
       y="Publications")

publications+theme_ADC()

```

```{r percent-graph, echo=FALSE}
percents<-ggplot()+
  geom_point(data = website_summary_metrics, aes(x=Date, y=Pub_Percent),color="#1D244E",size=2.5)+
  geom_point(data = website_summary_metrics, aes(x=Date, y=Data_Percent),color="#19B369",size=2.5)+
  geom_point(data = website_summary_metrics, aes(x=Date, y=Submit_Percent),color="#79FDB1",size=2.5)+
  labs(title='Percentage of Total Views by Page Type',
       x="Date",
       y="Percent (%)")


percents+theme_ADC()

```
